---
title: "Covid-19"
author: "Josh Haskell"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

***The data for this project was downloaded from the following website:
https://covidtracking.com/data/download***

```{r}
#Load libraries and data
library(dplyr)
library(ggplot2)
library(readr)

Covid_Data_All <- read_csv("Covid Data.csv") %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y"))
```
***I started by inspecting the data, and ran a few plots looking at the positivity and death rates over different time periods.***

```{r}
#Grouping by state and selecting columns
Covid_Data_State <- Covid_Data_All %>% 
  group_by(state)%>% 
  mutate(Total_Death = sum(deathIncrease), Total_Negative = sum(negativeIncrease), Total_Positive = sum(positiveIncrease), Total_Tests = Total_Negative + Total_Positive, Total_Positivity_Rate = Total_Positive/Total_Tests, Total_Death_Rate = Total_Death/Total_Positive) %>% 
  select(state,Total_Death, Total_Negative, Total_Positive, Total_Tests, Total_Positivity_Rate, Total_Death_Rate)

#Plot using all days
Covid_Data_State %>% 
  ggplot(aes(x = Total_Positivity_Rate, y = Total_Death_Rate))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(title = "Positivity Rate to Death Rate")

#Filtering out the early spring surge by using June 1st as the start date
Covid_Data_State_AfterMay <- Covid_Data_All %>% 
  group_by(state)%>% 
  filter(date >= "2020-6-01") %>% 
  mutate(Total_Death = sum(deathIncrease), Total_Negative = sum(negativeIncrease), Total_Positive = sum(positiveIncrease), Total_Tests = Total_Negative + Total_Positive, Total_Positivity_Rate = Total_Positive/Total_Tests, Total_Death_Rate = Total_Death/Total_Positive) %>% 
  select(date,state,Total_Death, Total_Negative, Total_Positive, Total_Tests, Total_Positivity_Rate, Total_Death_Rate)
  
#Plot using the June 1st through present 
Covid_Data_State_AfterMay %>% 
  ggplot(aes(x = Total_Positivity_Rate, y = Total_Death_Rate))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(title = "Positivity Rate to Death Rate Since June 1st")

#Using days since the beginning of September
Covid_Data_State_After_Sep <- Covid_Data_All %>% 
  group_by(state)%>% 
  filter(date >= "2020-9-01") %>% 
  mutate(Total_Death = sum(deathIncrease), Total_Negative = sum(negativeIncrease), Total_Positive = sum(positiveIncrease), Total_Tests = Total_Negative + Total_Positive, Total_Positivity_Rate = Total_Positive/Total_Tests, Total_Death_Rate = Total_Death/Total_Positive) %>% 
  select(date,state,Total_Death, Total_Negative, Total_Positive, Total_Tests, Total_Positivity_Rate, Total_Death_Rate)
  
#Plot using September 1st through present
Covid_Data_State_After_Sep %>% 
  ggplot(aes(x = Total_Positivity_Rate, y = Total_Death_Rate))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(title = "Positivity Rate to Death Rate Since September 1st")
```

***I looked at several different plots filtering for Missouri.***

```{r}
#Filter for MO
Covid_Data_MO <- Covid_Data_All %>% 
  filter(state == "MO") %>% 
  select(date, positiveIncrease,deathIncrease)
  
Covid_Data_MO %>% 
  ggplot(aes(x = date, y = positiveIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Positive Tests")

Covid_Data_MO %>% 
  ggplot(aes(x = date, y = deathIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Deaths")

#Filter for 30 days before Labor Day
Covid_Data_MO_BeforeLaborDay <- Covid_Data_All %>% 
  select(date,state,positiveIncrease,deathIncrease) %>% 
  filter(state == "MO", date >= "2020-08-07" & date < "2020-09-07") %>% 
  arrange(date)

Covid_Data_MO_BeforeLaborDay %>% 
  ggplot(aes(x = date,y = positiveIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Positive Tests 30 Days Before Labor Day")

Covid_Data_MO_BeforeLaborDay %>% 
  ggplot(aes(x = date,y = deathIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Deaths 30 Days Before Labor Day")

#Filter for 30 days after Labor Day
Covid_Data_MO_AfterLaborDay <- Covid_Data_All %>% 
  select(date,state,positiveIncrease,deathIncrease) %>% 
  filter(state == "MO", date >= "2020-09-07" & date <= "2020-10-07") %>% 
  arrange(date)

Covid_Data_MO_AfterLaborDay %>% 
  ggplot(aes(x = date,y = positiveIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Positive Tests 30 Days After Labor Day")

Covid_Data_MO_AfterLaborDay %>% 
  ggplot(aes(x = date,y = deathIncrease))+
  geom_smooth(se = FALSE)+
  labs(title = "MO Daily Deaths 30 Days After Labor Day")
```

## Question

***Did Labor Day significantly increase average daily Covid-19 cases in Missouri the 30 days following the holiday versus the 30 days prior? Stated formally:***

***Ho: Mu cases 30 days before = Mu cases 30 days after.***

***Ha: Mu cases 30 days before != Mu cases 30 days after.***

```{r}
#Paired t-test
Covid_Data_MO_BeforeLaborDay <- Covid_Data_All %>% 
  select(date,state,positiveIncrease) %>% 
  filter(state == "MO", date >= "2020-08-07" & date < "2020-09-07") %>% 
  arrange(date)

Covid_Data_MO_AfterLaborDay <- Covid_Data_All %>% 
  select(date,state,positiveIncrease) %>% 
  filter(state == "MO", date >= "2020-09-07" & date <= "2020-10-07") %>% 
  arrange(date)

#Descriptive Statistics
mean(Covid_Data_MO_BeforeLaborDay$positiveIncrease)
sd(Covid_Data_MO_BeforeLaborDay$positiveIncrease)
mean(Covid_Data_MO_AfterLaborDay$positiveIncrease)
sd(Covid_Data_MO_AfterLaborDay$positiveIncrease)

#Conducting Paired t-test
t.test(Covid_Data_MO_BeforeLaborDay$positiveIncrease, Covid_Data_MO_AfterLaborDay$positiveIncrease, paired = TRUE)
```
***With a p-value of 0.1489, which is greater than set alpha of 0.05, we retain the null hypothesis.***

***A paired t-test suggested Labor Day did not significantly increase the average number of daily cases (p = 0.1489) in MO.***

## Question

***My hypothesis question here is does the proportion of total Covid-19 deaths to total cases differ between Missouri and Wisconsin the 30 days after Labor Day?***

***Ho: Proportion MO = Proportion WI.***

***Ha: Proportion MO != Proportion WI.***

```{r}
#Two Sample test of proportions

#Total Deaths for MO
DeathsMO <- Covid_Data_All %>% 
  select(date,state,deathIncrease) %>% 
  filter(state == "MO", date > "2020-09-07" & date <= "2020-10-07") %>% 
  summarise(Deaths = sum(deathIncrease))

#Total Deaths for WI
DeathsWI <- Covid_Data_All %>% 
  select(date,state,deathIncrease) %>% 
  filter(state == "WI", date > "2020-09-07" & date <= "2020-10-07") %>% 
  summarise(Deaths = sum(deathIncrease))

#Total Cases for MO
TotalMO <- Covid_Data_All %>% 
  select(date,state,positiveIncrease) %>% 
  filter(state == "MO", date > "2020-09-07" & date <= "2020-10-07") %>% 
  summarise(Total_Cases = sum(positiveIncrease))

#Total Cases for WI
TotalWI <- Covid_Data_All %>% 
  select(date,state,positiveIncrease) %>% 
  filter(state == "WI", date > "2020-09-07" & date <= "2020-10-07") %>% 
  summarise(Total_Cases = sum(positiveIncrease))

#Conducting a two-sample test of proportions
died <- as.numeric(c(DeathsMO, DeathsWI))
total <- as.numeric(c(TotalMO, TotalWI))
prop.test(died, total)
```
***With a p-value of .0000, which is less than alpha of 0.05, we reject the null hypothesis.***

***A two-sample test of proportions suggested the death rate proportion for Missouri (prop = 0.014) and Wisconsin (prop = 0.004) for the 30 days following Labor Day were statistically different (p < 0.05).***



