---
title: "House Sale Price Model"
author: "Josh Haskell"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
```

# Train

***I will be using housing data to train a model for house sale price.***

## Pre-process
    
```{r}
#Load Data
house_data <- read_csv("train.csv")

#Removing variables with more than 10% missing observations, renaming and mutating variables
remove_variables <- which(colMeans(is.na(house_data))>0.1)
house_data_cleaned <- house_data[-remove_variables] %>% 
  rename(SecondFlrSF = '2ndFlrSF', FirstFlrSF = '1stFlrSF') %>% 
  mutate(TotalSF = SecondFlrSF + FirstFlrSF)
  
#Remaining missing values
#I assumed there was no masonry veneer for missing values, and changed values to None
house_data_cleaned$MasVnrType <- replace_na(house_data_cleaned$MasVnrType,  "None")
house_data_cleaned$MasVnrArea <- replace_na(house_data_cleaned$MasVnrArea,  "0")
  
#Since the majority (933 out of 1020) are Standard Circuit Breakers & Romex (SBrkr), I replaced any NAs to this value 
table(house_data_cleaned$Electrical)
house_data_cleaned$Electrical <- replace_na(house_data_cleaned$Electrical, "SBrkr")

#I used the average value for the year built
house_data_cleaned$GarageYrBlt <- replace_na(house_data_cleaned$GarageYrBlt, round(mean(house_data_cleaned$GarageYrBlt,na.rm = TRUE), digits = 0))

#For the rest of these values, I assumed there was no basement or garage, so I changed the value to None
house_data_cleaned$BsmtQual <- replace_na(house_data_cleaned$BsmtQual, "None")
house_data_cleaned$BsmtQual <- factor(house_data_cleaned$BsmtQual, levels = c("Ex","Gd","TA","Fa","Po","None"))

house_data_cleaned$BsmtCond <- replace_na(house_data_cleaned$BsmtCond, "None")
house_data_cleaned$BsmtCond <- factor(house_data_cleaned$BsmtCond, levels = c("Ex","Gd","TA","Fa","Po","None"))

house_data_cleaned$BsmtExposure <- replace_na(house_data_cleaned$BsmtExposure, "None")
house_data_cleaned$BsmtExposure <- factor(house_data_cleaned$BsmtExposure,levels = c("Gd","Av","Mn","No","None"))

house_data_cleaned$BsmtFinType1 <- replace_na(house_data_cleaned$BsmtFinType1, "None") 
house_data_cleaned$BsmtFinType1 <- factor(house_data_cleaned$BsmtFinType1,levels = c("GLQ","ALQ","BLQ","Rec","LwQ","Unf","None"))

house_data_cleaned$BsmtFinType2 <- replace_na(house_data_cleaned$BsmtFinType2, "None")
house_data_cleaned$BsmtFinType2 <- factor(house_data_cleaned$BsmtFinType2, levels = c("GLQ","ALQ","BLQ","Rec","LwQ","Unf","None"))

house_data_cleaned$GarageType <- replace_na(house_data_cleaned$GarageType, "None")
house_data_cleaned$GarageType <- factor(house_data_cleaned$GarageType,levels = c("2Types","Attchd","Basment","BuiltIn","CarPort","Detchd","None"))

house_data_cleaned$GarageFinish <- replace_na(house_data_cleaned$GarageFinish, "None")
house_data_cleaned$GarageFinish <- factor(house_data_cleaned$GarageFinish,levels = c("Fin","RFn","Unf","None"))

house_data_cleaned$GarageQual <- replace_na(house_data_cleaned$GarageQual, "None")
house_data_cleaned$GarageQual <- factor(house_data_cleaned$GarageQual,levels = c("Ex","Gd","TA","Fa","Po","None"))

house_data_cleaned$GarageCond <- replace_na(house_data_cleaned$GarageCond, "None")
house_data_cleaned$GarageCond <- factor(house_data_cleaned$GarageCond, levels = c("Ex","Gd","TA","Fa","Po","None"))
```

## Model

***Modeling sale price as a function of chosen variables.***

```{r}
#inspecting different variables
house_data_cleaned %>% 
  select(SalePrice,LotArea,BedroomAbvGr,YrSold, FullBath,HalfBath,GarageCars,TotalSF, PoolArea) %>% 
  pivot_longer(-SalePrice,names_to = "Variable", values_to = "Value") %>% 
  ggplot(aes(x = Value, SalePrice))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~Variable, scales = "free_x")

#Removing outliers for Lot Area
house_data_cleaned <- house_data_cleaned %>% 
  filter(LotArea < 30000)

#Removing outliers for TotalSF
house_data_cleaned <- house_data_cleaned %>% 
  filter(TotalSF < 3500) 
```
```{r}
#taking log of SalePrice
house_data_cleaned %>% 
  select(SalePrice,LotArea,BedroomAbvGr,YrSold, FullBath,HalfBath,GarageCars,TotalSF, PoolArea) %>% 
  pivot_longer(-SalePrice,names_to = "Variable", values_to = "Value") %>% 
  ggplot(aes(x = Value, log(SalePrice)))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~Variable, scales = "free_x")
```

***I decided to leave PoolArea out of the final model due to outliers. The year sold had very little linear relationship, so I left that variable out too. I decided to take the log of sale price since the data points were heteroskedastic.***


```{r}
mymodel <- lm(log(SalePrice) ~ LotArea + BedroomAbvGr + Neighborhood + FullBath + HalfBath + GarageCars + TotalSF, data = house_data_cleaned)

mean(residuals(mymodel))

summary(mymodel)
```
***With a p-value and residuals both near zero, this would suggest the model is a good fit. R-squared implies 81% of variability in the house sale price is explained by these variables.***

# Test

## Predict

***I will be using test data to predict values from the model.***

```{r}
test_data <- read_csv("test.csv")

remove_test_variables <- which(colMeans(is.na(test_data))>0.1)
test_data_cleaned <- house_data[-remove_test_variables] %>% 
  rename(SecondFlrSF = '2ndFlrSF', FirstFlrSF = '1stFlrSF') %>% 
  mutate(TotalSF = SecondFlrSF + FirstFlrSF)

test_data_prediction <- as.data.frame(predict(mymodel, test_data_cleaned, interval = "prediction"))
```

## Evaluate

***Using predicted values from above to calculate the root mean square error of the predictions on the test data. This is one way to measure how well the model fits the test data.***

```{r}
RMSE_Test_Data <- sqrt((1/length(test_data_prediction$fit))*sum((test_data_prediction$fit - mean(test_data_prediction$fit))^2))
RMSE_Test_Data
```
***With RMSE of the predicted values of 0.36, compared to 0.17 for the existing model, it suggests a good fit.***